---
title: "Report"
author: "ADS2 Group 4"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  pdf_document:
    df_print: tibble
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
    toc: yes
    toc_depth: 5
  word_document:
    toc: yes
    toc_depth: '5'
geometry: inner=0.5in,outer=0.5in
---


# Global Libraries

Locating global libraries.

```{r Global libraries}
# Remove cached variables
rm(list = ls())
# Knitr Options
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
# Check whether the dependencies are met
installed <- installed.packages()[, 1]
required <- c("tidyverse", "ggubr", "knitr", "rmarkdown", "stringr", "viridis", "C50", "pheatmap")
for (packages in required) {
  if (!packages %in% installed) {
    simpleError(paste0("Package ", packages, " not installed."))
  }
}
rm(installed)
rm(packages)
# Source global packages
library(tidyverse)
library(C50)
library(ggpubr)
library(viridis)
library(pheatmap)
```

# Data Cleaning

```{r Read Data}
if (!file.exists("substance_use.csv")) simpleError(paste0("File substance_use.csv not exist!"))
sub_use <- read.csv("substance_use.csv", stringsAsFactors = TRUE)
```

Use `head` and `summary` to explore data.

```{r Summary}
summary(sub_use)
slice_head(sub_use, n = 3)
```

We discovered that there is only `"Percent"` in the `metric` column. This column is therefore removed.

```{r Remove metric}
summary(factor(sub_use$metric))
sub_use <- select(sub_use, !(metric))
```

Check whether there are `NA`s inside.

```{r deNA}
summary(complete.cases(sub_use))
```

It is clear that there is no `NA` inside.

# Part 1: Exploring the Data

## Question 1: Which region of the world has the highest rate of alcohol-related deaths among men aged 40-44 in 2019?

From the question, we can conclude that the data group required is the men between the ages of 40 and 44 who died of alcohol use disorder in 2019. We get all data satisfying the criteria and then sort the data by val in reverse order. Here "val" refers to the death rate. The data with the highest death rate is placed at the top. We pick the highest one and get its location.

```{r P1Q1}
# Get all data satisfying the criteria
sub_use40_44 <- sub_use[sub_use$year == 2019 &
                          sub_use$sex == "Male" &
                          sub_use$cause == "Alcohol use disorders" &
                          sub_use$age == "40 to 44" &
                          sub_use$measure == "Deaths",]
# Sort the result by val in reverse order
sub_use_arranged <- arrange(sub_use40_44, desc(val))
# Pick the highest one and get its location
sub_use_arranged[1, "location"]
rm(sub_use40_44)
rm(sub_use_arranged)
```

We conclude that Europe and Central Asia -WB have the highest rate of alcohol-related deaths among men aged 40-44 in 2019.

## Question 2: Looking at the prevalence of alcohol-related disease in the East Asia and Pacific region, how has this changed over time and in the different age groups? Is there a difference between men and women?

From the question, we get the criteria that measure equals to "prevalence", location equals to "East Asia" or "Pacific -WB" and cause equals to "Alcohol use disorders". We get all data satisfying the criteria. We use ggplot function to plot scatter plots of the data. X-axis displays the "year" and Y-axis displays the "val". Age and sex are distinguished by color and shape. We facet the plot according to sex so that we can clearly see the difference between male and female. 

```{r P1Q2.1}
sub_use_alcohol <- sub_use[sub_use$measure == "Prevalence" &
                             sub_use$location == "East Asia & Pacific - WB" &
                             sub_use$cause == "Alcohol use disorders",]

g <- ggplot(sub_use_alcohol, aes(x = year, y = val, color = age, shape = sex)) +
  geom_point(show.legend = TRUE) +
  geom_line(show.legend = TRUE) +
  facet_grid(. ~ sex) +
  labs(title = "Prevalence of alcohol-related disease in the East Asia and Pacific region",
       x = "Year", y = "Prevalence") +
  scale_color_viridis(discrete = T)+
  theme(panel.spacing.x = unit(1.2,"lines"))
g
```

From the graph, we can know that:

* For males, the prevalence of alcohol-related disease in the East Asia and Pacific region rises from 1990 to 1995, and decreases to 2005, and then rise again non-stop for those aged 25 to 54. For those who are 55 to 69, they reached to peak in about 2017 and decreases afterwards.
* Comparing to males, there are no significant changes in females.

To determine whether there is a difference between man and woman, we plot a violin plot as statistical tests are explicitly announced to be unnecessary.

```{r P1Q2.2}
g <- ggplot(sub_use_alcohol) +
  geom_violin(aes(x = sex, y = val, fill = sex), show.legend = TRUE) +
  labs(title = "Prevalence of alcohol-related disease in the East Asia and Pacific region",
       x = "Gender", y = "Prevalence") +
  scale_fill_manual(values = c("#E69F00", "#56B4E9"))
g
rm(sub_use_alcohol)
rm(g)
```

From this violin plot, we can conclude that there is a difference between male and female. Female is significantly lower than male.

## Question 3: Looking at the data from the United States, can you confirm an increase in the prevalence of diseases related to opioid use? What age group is the most affected?

The remaining part of the description is:

> In the United States, there is talk of an "Opioid epidemic". Part of the problem is that since the late 1990s, doctors have increasingly been prescribing pain killers which can be highly addictive.

From the question, we get the criteria that measure equals to "Prevalence", location equals to "North America" and cause equals to "Opioid use disorders". We get all data satisfying the criteria. We use ggplot function to plot scatter plots of the data. X-axis displays the "year" and Y-axis displays the "val". Age is distinguished by color. We facet the plot according to sex so that we can see the difference between male and female. 

```{r P1Q3}
sub_use_Opioid <- sub_use[sub_use$measure == "Prevalence" &
                            sub_use$location == "North America" &
                            sub_use$cause == "Opioid use disorders",]

g <- ggplot(sub_use_Opioid, aes(x = year, y = val, color = age)) +
  geom_point(show.legend = TRUE) +
  geom_line(show.legend = TRUE) +
  labs(title = "Prevalence of opioid-related disease in the United States",
       x = "Year", y = "Prevalence") +
  facet_grid(. ~ sex) +
  scale_color_viridis(discrete = T)+
  theme(panel.spacing.x = unit(1.2,"lines"))
g
rm(sub_use_Opioid)
rm(g)
```

From the plot, we can confirm an increase in the prevalence of diseases related to opioid use. Both men and women in 25-29 age group are the most affected because of their highest initial values and growth of the opioid-related prevalence.

# Part 2: Ask your own Question

## Question 4: When we plot the "val" according to "year" we can find that there are some visible differences among locations. Therefore it leads to our own question: Are there any rules for distinguishing the locations by other attributes?

In this part, we will build a classification tree model using C5.0 algorithm provided by R package `C50`. The following part reveals the training process, and the question we ask will be stated below.


First, let us take the opioid-related prevalence as an example. We use ggplot function to plot scatter plots of the prevalence value of Opioid use disorders and facet it according to sex and location. 

```{r,fig.width=8, fig.height=8}

sub_use_Opioid <- sub_use[sub_use$measure == "Prevalence" &
                                 sub_use$cause == "Opioid use disorders",]

g <- ggplot(sub_use_Opioid, aes(x = year, y = val, color = age)) +
        geom_point(show.legend = TRUE) +
        geom_line(show.legend = TRUE) +
        labs(title = "Prevalence of opioid-related disease in all locations",
             x = "Year", y = "Prevalence") +
        #facet_grid(location ~ sex) +
        scale_color_viridis(discrete = T)+
  theme(panel.spacing.x = unit(1.2,"lines"))
g<-facet(g,facet.by = c("location","sex"),panel.labs.font.y = list(angle=0),scales = "free")
g
```

As can be seen from the above figure, some visible differences can be seen among other regions. Therefore, we decided to explore whether the distinction between locations exist and the contribution of each factor (sex, age, cause, val, upper, lower, measure, year) to the region, that is, how much difference each data contributes.

```{r Training}
sub_use_rand <- sub_use[order(runif(15000)),]
train <- slice_head(sub_use_rand, n = 15000 / 2)
train.data <- select(train, !(location))
train.label <- train$location
test <- slice_tail(sub_use_rand, n = 15000 / 2)
test.data <- select(test, !(location))
test.label <- test$location
test.location <- unique(as.character(test.label))

# Training the model
m <- C5.0(train.data, train.label, trials = 100)

rm(sub_use_rand)
rm(train)
rm(train.data)
rm(train.label)
rm(test)
```

The following table shows factors that do matter in this module.

```{r Factors}
#Print the usage attribute
attributeUsage <- strsplit(m$output, "Attribute usage")[[1]][2]
attributeUsage<-paste("Attribute usage", attributeUsage,sep="")
cat(attributeUsage)
rm(attributeUsage)
```
From the table, we can see that age and cause have a greater impact on the region, followed by year and gender. The contribution of measurement indicators is the smallest, which is in line with our common sense.

The following graph shows the error rate, from which we can know that this model is over-fitting.

```{r Error Rate}
err <- m$boostResults
ggscatter(err, x = "Trial", y = "Percent", add = "reg.line",
          add.params = list(color = "blue", fill = "lightgray"), conf.int = TRUE) +
  geom_line()
rm(err)
```

The following piece of code performs the prediction.

```{r Prediction}
pred <- predict(m, test.data, type = "class")
rm(test.data)
```

The following graph shows the confusion matrix, both unnormalized and normalized with Log-Normalize.

```{r Confusion Matrix}
pred <- as.character(pred)
test.label <- as.character(test.label)
df <- as.data.frame(matrix(0,
                           nrow = length(test.location),
                           ncol = length(test.location)),
                    row.names = test.location)
names(df) <- test.location
for (i in seq(1, length(pred))) {
  df[pred[i], test.label[i]] <- df[pred[i], test.label[i]] + 1
}
pheatmap(df, cluster_cols = FALSE, cluster_rows = FALSE,
         display_numbers = TRUE,angle_col = 315)
# log normalize
pheatmap(log10(df + 1), cluster_cols = FALSE, cluster_rows = FALSE,
         display_numbers = TRUE,angle_col = 315)
rm(df)
rm(i)
rm(test.label)
rm(test.location)
```

It can be found from the two heat maps that the predicted result is close to the actual result, which proves that our machine learning model is basically correct. Additionally, the existence of a practical decision tree can confirm the significant differences between locations.



# Cleaning Up


```{r Cleaning Up}
rm(m)
rm(pred)
rm(sub_use)
```

# Supplementary Code

This RMarkdown should be compiled with following code:

```{r Compile, eval = FALSE}
library(rmarkdown)
library(knitr)
eng_text <- function(options) {
  engine_output(options, code = options$code, out = "")
}
knit_engines$set(text = eng_text)
render("Main.Rmd", output_format = "all")
```
